### 2023-05-01 chapter2/chapter2.tex
Removed outline for discussion paragraph on PCAs
Updated text with examples of clusters

\subsection{A model of constrained evolution reveals conserved features}
Though the BM process permits the inference of the rates of feature evolution after accounting for the phylogenetic relationships between species, it does not directly test for their conservation. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time. Instead, evidence of conservation requires comparison to a model where trait variation is constrained. A common choice for modeling the effect of selection on the evolution of a continuous trait is the Ornstein-Uhlenbeck (OU) model. The OU model is similar to the BM model where a trait accumulates variation through small, undirected steps. However, it differs in the trait is also attracted towards an optimal value where the attraction is proportional to a parameter $\alpha$ and the trait's distance from this value. Under an additional assumption of stationarity that ensures parameter identifiability and estimate consistency, the OU model is specified with three total parameters: the optimal value, the fluctuation magnitude, and the selection strength~\cite{Ho2013, Ho2014}. While the first two parameters are analogous to the root and rate parameters in the BM model, respectively, the selection strength has no equivalent.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hierarchy/out/hierarchy.png}
\centering
\caption{\textbf{Hierarchical clustering of evolutionary signatures.}
The AIC difference between the BM and OU models is measure of their relative goodness of fit to the data with a penalty for the number of parameters in each. Larger values indicate a better fit by the OU model. Clusters are indicated by rectangles on the right. The cluster descriptions are summaries of their enriched annotations and do not necessarily correspond to specific terms.}
\label{fig:hierarchy}
\end{figure}

The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise (Fig.~\ref{fig:hierarchy}). However, we identify at least 12 clusters with a strong and consistent patterns of constraint, with many having significant associations with certain processes, functions, or compartments. For example, cluster X is strongly enriched in Y, and cluster A contains many proteins involved in B. Though these preliminary results suggest the existence of high-level relationships between certain patterns of conserved features and function, further insight into the biophysical and mechanistic details that underpin these associations will rely on more in-depth analyses of specific clusters and the properties of their constituent IDRs.

@@

% IV. Feature analyses
%    A. Though no discrete clusters are visible in the largest two and three principal components of the feature root and rate distribution, they only account for X\% and Y\% of the total variation between regions, respectively. Thus, a majority of the variation is contained in other components.
%    B. The success of the hierarchical clustering at identifying common patterns of feature constraint suggests this method, or other non-linear dimensionality reductions like UMAP or t-SNE, may also reveal additional structure latent in the high-dimesional rate distributions. However, as these methods are designed primarily for visualization, they can significant distort the global structure of high-dimensional data. Thus, these embeddings are generally not suitable for downstream analyses. Instead, semi-supervised approaches for feature learning may yield more meaningful representations. In fact, a recent study used a semi-supervised learning approach that identified orthologous IDRs to learn novel disorder-associated features. Though they used a simple neural network that aided in interpretability, other similar semi-supervised approaches have used more complex networks to learn rich representations of protein regions. Similar approaches may yield representations of the space of disordered proteins that are more clearly visualized and manipulated.

>>>>

\subsection{A model of constrained evolution reveals patterns of conserved features}
Though the BM process permits the inference of the rates of feature evolution after accounting for the phylogenetic relationships between species, it does not directly test for their conservation. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time. Instead, evidence of conservation requires comparison to a model where trait variation is constrained. A common choice for modeling the effect of selection on the evolution of a continuous trait is the Ornstein-Uhlenbeck (OU) model. The OU model is similar to the BM model where a trait accumulates variation through small, undirected steps. However, it differs in the trait is also attracted towards an optimal value where the attraction is proportional to a parameter $\alpha$ and the trait's distance from this value. Under an additional assumption of stationarity that ensures parameter identifiability and estimate consistency, the OU model is specified with three total parameters: the optimal value, the fluctuation magnitude, and the selection strength~\cite{Ho2013, Ho2014}. While the first two parameters are analogous to the root and rate parameters in the BM model, respectively, the selection strength has no equivalent.

@@

\begin{figure}[h!]
\includegraphics[width=\textwidth]{hierarchy/out/hierarchy.png}
\centering
\caption{\textbf{Hierarchical clustering of evolutionary signatures.}
The AIC difference between the BM and OU models is measure of their relative goodness of fit to the data with a penalty for the number of parameters in each. Larger values indicate a better fit by the OU model. Clusters are indicated by rectangles on the right. The cluster descriptions are summaries of selected significantly enriched annotations ($p < 0.01$) and do not necessarily correspond to specific terms.}
\label{fig:hierarchy}
\end{figure}

The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise (Fig.~\ref{fig:hierarchy}). However, we identify at least 12 clusters with a strong and consistent patterns of constraint, with many having significant associations with certain processes, functions, or compartments. For example, cluster B is strongly enriched in proteins associated with the Wnt signaling pathway, and cluster F contains many proteins involved in nuclear transport. (Selected annotations corresponding to the summaries in Fig.~\ref{fig:hierarchy} are given in Table~\ref{stable:annotations}.)  Though these preliminary results suggest the existence of high-level relationships between certain patterns of conserved features and function, further insight into the biophysical and mechanistic details that underpin these associations will rely on more in-depth analyses of specific clusters and the properties of their constituent IDRs.

@@
